Structural replication of nanoporous media using procedural noise

### 2017-02-05 ms.tex
Wrote more about g(r), lammps simulation

\documentclass[aps,pre,twocolumn,letterpaper,floatfix,showpacs]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\usepackage{color}
\definecolor{Blue}{rgb}{0.1,0.1,1.0} 
\definecolor{Magenta}{rgb}{1.0,0.1,0.5} 
\definecolor{LRed}{rgb}{0.8,0.0,0.0}

@@

We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations. Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the pair correlation function $g(r)$. 
\todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
%\section{Method}
\section{Procedural content generation}
By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns. A firm example would be fractals, such as the Julaia/Mandelbrot [ref mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane. With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
\begin{figure}
\includegraphics[width=.5\textwidth]{orphancity.png}
\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods. Created by one of the authors. \todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
\label{fig:orphancity}
\end{figure}

@@

\section{Model}
In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media. For a given data set and a model, we wish to estimate the parameters that best represent the data. This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system. The simulated system is confined to a cube of size $159 \AA^3$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.

@@

Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest. For instance, pore size distribution and porosity are import for studying transport in porous media. In this paper, we choose to use the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}. The $g(r)$ (or pair correlation function) describes how density varies as a function of distance from a reference particle. Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ are the set of model parameters.

The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the triangles.

@@

\section{Results}
We now turn our attention to applying the MCMC framework on two kinds of data. First, we wish to test and verify the code by performing a full data analysis on mock data with known input parameters. This enables us to investigate the posterior distribution, and determine whether any potential problems could arise with the likelihood landscape, such as radom walkers trapped in local minima. We then continue with a full analysis of simulated SiO2 data, and see whether the resulting best-fit parameter data set is consistent with the SiO2 data set.

@@

\subsection{Parameter estimation of simulated SiO2 silica}
%\begin{figure}
%\includegraphics[width=.49\textwidth]{sio_porous.png}
%\caption{Simulated silica created with LAMMPS. The box is $159^3 \AA$, with $100 000$ particles, having been created by cooling an expanding system until quenching was achieved. Note the similarities between the pores here and in figure \ref{fig:mockdata}.}
%\label{fig:si2data}
%\end{figure}
\begin{figure*}
\includegraphics[width=.99\textwidth]{results_porous_full.png}
\caption{Results from the analysis of a simulated $159^3$ \AA SiO2 nanoporous data set. The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors. Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.}
\label{fig:porous_results1}
\end{figure*}

We now perform a parameter fitting analysis on realistic simulated nanoporous media. Using LAMMPS \cite{plimpton1995fast}, we simulate a $100^3\AA$  box of 100K SiO2 silica particles at an initial temperature of $4500 K$, using the Vashishta-potential. After an initial period, we expand the box to $159^3 \AA$ and start cooling until quenching is achieved at about $300K$. The resulting nanoporous media is depicted on the right side in figure \ref{fig:porous_vs_model}, with a porosity of $75\%$. \todo{Skrive mer om lagingen?}

With the model described in equation \ref{eq:noisemodel1}, we perform a full likelihood analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure. The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, with best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood. We also display the RFD $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrated the sensitivity of $g(r)$ given the model parameters. Note that we are not able to accurately reproduce the structures of the SiO2 $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.

\begin{figure}
\includegraphics[width=.45\textwidth]{gofr_plot.png}
\caption{The RDF $g(r)$ for three data sets: the simulated SiO2 input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green). }
\label{fig:gofr1}
\end{figure}
Finally, we depict a comparison of the actual simulated input SiO2 data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
\begin{figure}
\includegraphics[width=.45\textwidth]{comparison.png}
\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO2 input data set used for parameter estimation. }
\label{fig:porous_vs_model}
\end{figure}

@@

\subsection{$g(r)$ variation}
We also calculated the variation in the RDF $g(r)$ given initial conditions on both our model and simulated data. In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMBS simulation. For the model, it corresponds to the random seed. We found that varying these initial conditions have negligible effect on the calculation of $g(r)$.

@@

\section{Conclusion}
\todo{konklusjon maa renskrives}
In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise. We have developed tools for creating such porous materials, and also estimating model parameters given data. We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model. In addition, we performed an analysis on a simulated $159^3 \AA$ data set of SiO2, and explain that while porosity and surface area corresponds well between the model and data, the RDF $g(r)$ shows that the current model does not exactly reproduce correct data. However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.

>>>>

\documentclass[aps,pre,twocolumn,letterpaper,floatfix,showpacs]{revtex4}
\usepackage{graphicx} 
\usepackage{amsmath,amssymb,amsfonts} 
\usepackage{mathtools}
\usepackage{pdfpages}
\usepackage{afterpage}
\usepackage[hidelinks]{hyperref} 
\usepackage{epstopdf}
\usepackage{siunitx}
\usepackage{color}
\definecolor{Blue}{rgb}{0.1,0.1,1.0} 
\definecolor{Magenta}{rgb}{1.0,0.1,0.5} 
\definecolor{LRed}{rgb}{0.8,0.0,0.0}

@@

We do not claim that procedural noise methods completely reproduces all structural properties of nanoporous silica, but show that our approach can be used to efficiently create materials that are structurally similar to simulations. Naturally, this all depends on the statistics used for comparing the two models, which in this case is based on the radial distribution function $g(r)$ (RDF). 
\todo{Nevne at kjemi naturligvis ikke vil være riktig siden antall naboer per atom ikke er riktig. }
%\section{Method}
\section{Procedural content generation}
By procedural content generation, we generally mean a mathematical algorithm that uses limited input data to create complex, sometimes infinite repeating patterns. A firm example would be fractals, such as the Julaia/Mandelbrot [ref mandelbrot] set, where one tests whether the complex quadratic recurrence equation $z_{n+1} = z_n^2 +C$ for $z \in \mathbb C$ converges for each position in a 2D complex plane. With this seemingly simple equation, the level of complexity generated on the boundary between convergence and divergence is infinite, with self-similar patterns repeating themselves to arbitrary detail.
\begin{figure}
\includegraphics[width=.5\textwidth]{orphancity.png}
\caption{Example of city building using procedural methods: Landscape, city infrastructure, buildings and all textures are all automatically generated using procedural methods. Created by one of the authors. \todo{Denne er jo med bare for moro skyld, boer jo fjernes} }
\label{fig:orphancity}
\end{figure}

@@

\section{Model}
In this paper, we are interested in developing a model using simplex noise to reproduce nanoporous structures that are as statistically close to actual simulations of porous media. For a given data set and a model, we wish to estimate the parameters that best represent the data. This is performed through a full likelihood analysis, where we in addition validate the model by comparing the radial distribution function $g(r)$, porosity and surface area of the system. The simulated system is confined to a cube of size $\SI{159}{\angstrom^{3}}$, but since the procedural noise function is continuous on all scales, the box could be of arbitrary size.

@@

Ideally, the model measure should pick up all geometrical properties that are relevant for the physical study of interest. For instance, transport in porous media heavily depends on the pore size distribution and porosity, whereas chemical reactons are sensitive to surface area. \todo{referanser, malthe?} In this paper, we choose to focus on the radial distribution function $g(r)$ as defined and implemented in LAMMPS \cite{plimpton1995fast}. $g(r)$ describes how density varies as a function of distance from a reference particle, picking up both the intramolecular and intermolecular structure. We have chosen to ignore the short range contributions of $g(r)$ since the pore geometry does not depend on the intramolecular structure of SiO$_2$, but only the density variations at larger distances. Hereafter, $g_d(r)$ denotes the RDF calculated from simulated or experimental data, while $g_{M(\bar \theta)}(r)$ represents the RDF based on the procedural noise model $M(\bar \theta)$, where $\bar \theta$ are the set of model parameters.

The method used for calculating porosity of a media is defined as in \cite{gelb1998characterization}, and we obtain the surface area of a porous media by using marching cubes triangulation to sum the surface area of the generated triangles. As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in atomic simulations, because a definition of pore space is needed. In the marching cubes triangulation, we generate a voxelized grid, the scalar field, with values being the distance from the voxel center to the nearest atom. The isovalue threshold defining the surface is chosen to be \SI{2}{\angstrom}, large nough to have a dense pore matrix, but small enough to have most of the available pore space defined as such. The important note here is that the same definition of surface is being used in all compared systems.

@@

\section{Results}
We now turn our attention to applying the MCMC framework on two kinds of data. First, we wish to test and verify the code by performing a full data analysis on mock data with known input parameters. This enables us to investigate the posterior distribution, and determine whether any potential problems could arise with the likelihood landscape, such as radom walkers trapped in local minima. We then continue with a full analysis of simulated SiO$_2$ data, and see whether the resulting best-fit parameter data set is consistent with the SiO$_2$ data set.

@@

\subsection{Parameter estimation of simulated nanoporous silica}
%\begin{figure}
%\includegraphics[width=.49\textwidth]{sio_porous.png}
%\caption{Simulated silica created with LAMMPS. The box is $159^3 \AA$, with $100 000$ particles, having been created by cooling an expanding system until quenching was achieved. Note the similarities between the pores here and in figure \ref{fig:mockdata}.}
%\label{fig:si2data}
%\end{figure}
\begin{figure*}
\includegraphics[width=.99\textwidth]{results_porous_full.png}
\caption{Results from the analysis of a simulated $159^3$ \AA SiO$_2$ nanoporous data set. The upper panel displays the marginalised 1D posteriors for each parameter, while the lower panel depicts the marginalised 2D posteriors. Best fit parameters from the maximum 3D likelihood suggest $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$.}
\label{fig:porous_results1}
\end{figure*}

We now perform a parameter fitting analysis on realistic simulated nanoporous media. Using LAMMPS \cite{plimpton1995fast} and the Vashishta-potential \cite{vashishta1990interaction}, we simulate system consisting of $14^3$ betacristobalite unit cells of SiO$_2$, corresponding to system size of approximately \SI{100}{\angstrom}. The temperature is then increased to \SI{4500}{\kelvin} using NVT integration. After an initial period, we expand the box to \SI{159}{\angstrom} to achieve a porosity $\phi=0.75$. Then, the system is quenched by setting the temperature to \SI{300}{\kelvin}, resulting in the nanoporous media on the right side in figure \ref{fig:porous_vs_model}. By choosing a different seed for the initial velocities, it is easy to produce other samples.

With the model described in equation \ref{eq:noisemodel1}, we perform a full likelihood analysis of the data set with the three free parameters $\tau$, $\nu$ and $\psi$ using $g(r)$ as the measure. The resulting marginalised posteriors are depicted in figure \ref{fig:porous_results1}, with best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$ is extracted from the full 3D likelihood. We also display $g(r)$ for the input data set, a best-fit parameter data set and a bad-fit data set in figure \ref{fig:gofr1}, to illustrate the sensitivity of $g(r)$ given the model parameters. Note that we are not able to accurately reproduce the structures of the SiO$_2$ $g(r)$ with our current model, but hope that we during future work will be able to produce another model that provides a better fit with this measure.

Note that the surface atoms are not in chemical equilibrium since some neighbors of a surface atom might have been removed. In order to use the generated geometry in molecular dynamics, one would need either noble gas atoms which may be stable, or apply passivation step.

\begin{figure}
\includegraphics[width=.45\textwidth]{gofr_plot.png}
\caption{$g(r)$ for three data sets: the simulated SiO$_2$ input data set (red), the best-fit parameter model (blue) and a random low-fit parameter model (green). \todo{Bytt denne figuren med VMD sin beregning. Normaliseringen vår er gal.}}
\label{fig:gofr1}
\end{figure}
Finally, we depict a comparison of the actual simulated input SiO$_2$ data set together with a representation of the best-fit parameters in figure \ref{fig:porous_vs_model}   
\begin{figure}
\includegraphics[width=.45\textwidth]{comparison.png}
\caption{Right side: A realisation of the best-fit parameters $\tau=-0.2$, $\psi=0.8$ and $\nu=0.3$. Left side: Simulated SiO$_2$ input data set used for parameter estimation. }
\label{fig:porous_vs_model}
\end{figure}

@@

\subsection{$g(r)$ variation}
We also calculated the variation in the $g(r)$ given initial conditions on both our model and simulated data. In terms of the data, the variation is attributed as the initial velocities of the particles in the LAMMPS simulation. For the model, it corresponds to the random seed. We found that varying these initial conditions have negligible effect on the calculation of $g(r)$.

@@

\section{Conclusion}
\todo{konklusjon maa renskrives}
In this paper, we have proposed a novel method for generating a simple kind of nanoporous media using procedural methods, namely simplex noise. We have developed tools for creating such porous materials, and also estimating model parameters given data. We have shown that it is possible through a full likelihood analysis to reproduce the input parameters of a simulated model. In addition, we performed an analysis on a simulated $159^3 \AA$ data set of SiO$_2$, and explain that while porosity and surface area corresponds well between the model and data, the $g(r)$ shows that the current model does not exactly reproduce correct data. However, we do not claim that the current model should do so yet, since we have chosen the simplest kind of summation with regards to procedural noise modes.
